---
title: "xgboost_numero1"
author: "Vera"
date: "14 7 2017"
output: html_document
---

Libraries:
```{r, echo = F}
library(data.table)
library(dplyr)
library(tidyr)
library(xgboost)
library(ggplot2)

vh <- function(x){View(head(x,25))}
```

### Prep: 

Loading Data:
* Aisles and Department data just contain the name belonging to the number; not needed here as xgboost needs hot encoded or label encoded(numeric) features. 
* op_prior and op_train contain product information from training and test people (! train and test people are not the same ones) -> features: order_id, product_id, add_to_cart_order, reordered(1,2)
* orders contains some information about an order for train, test and prior orders. With this dataset test useres can be identified. Features: order_id, user_id, eval_set, order_number, dow, hours and days since last order.
* products contains informarion about the product -> aisle and department information

```{r}
aisles <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/aisles.csv.zip", "aisles.csv"), stringsAsFactors = FALSE)

departments <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/departments.csv.zip", "departments.csv"), stringsAsFactors = FALSE)

op_prior <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/order_products__prior.csv.zip", "order_products__prior.csv"), stringsAsFactors = FALSE)

op_train <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/order_products__train.csv.zip", "order_products__train.csv"), stringsAsFactors = FALSE)

orders <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/orders.csv.zip", "orders.csv"), stringsAsFactors = FALSE)
#vh(orders)

products <- read.csv(unz("/home/Vera_Weidmann/Supermarket/00_Data/products.csv.zip", "products.csv"), stringsAsFactors = FALSE)
```

Reshape data:
```{r}
# merge user_id to op_train and op_prior
op_train$user_id <- orders$user_id[match(op_train$order_id, orders$order_id)]
op_prior$user_id <- orders$user_id[match(op_prior$order_id, orders$order_id)]

# merge the information about orders to prior order behavior. This data contains just information about prior orders form training and test users. 
orders_products <- orders %>% inner_join(op_prior, by = c("order_id", "user_id"))
head(orders_products)

#These are the test users and orders (which are contained by the dataset orders)
testorders <- orders %>% filter(eval_set=="test") %>% select(user_id,order_id)
testusers <- testorders[,1] #75000 users
```

### Product Features:

Here we create new features which decribes the product, based on the prior order behavior:
(* product_time: How often a product occured based on the user order_history)
* prod_orders: How often a product was bought over all users and baskets
* prod_reorderes: How often a product was reordered over all users and baskets
* prod_first_orders: reflects how often a product was added the first time to a basket
* prod_second_orders: reflects how often a product was added the second time to a basket
* prod_reorder_probability: How often a product was added a second time
* prod_reorder_times: probability of how often a product was reordered from all products which were bought to first time.
* prod_reorder_ratio: probaility of how often a product was reordered from all product purchases. 

```{r}

prd <- orders_products %>% #just prior information
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup() %>%
  group_by(product_id) %>%
  summarise(
    prod_orders = n(), #count products
    prod_reorders = sum(reordered), #count products reordered
    prod_first_orders = sum(product_time == 1), #product bought first time
    prod_second_orders = sum(product_time == 2) #product bought second time or more often
  )

prd$prod_reorder_probability <- prd$prod_second_orders / prd$prod_first_orders
prd$prod_reorder_times <- 1 + prd$prod_reorders / prd$prod_first_orders
prd$prod_reorder_ratio <- prd$prod_reorders / prd$prod_orders

#joining aisles and depertment ids
prd <- prd %>%
  inner_join(products, by = "product_id")

# chose features you want to keep
#prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)
prd <- prd %>% select(-product_name)

vh(prd)
```


### User Features:

These features also based on the prior user behavior. These features are created for training and test users.
* user_orders: how many orderes a user has
* user_period: how many days lie between the first order and the last one. 
* user_mean_days_since_prior: average time between orders.
* user_total_products: how many products had the user already bought
* user_reorder_ratio: how many products are reordered 
* user_distinct_products: how many unique products does the user buy
* user_average_basket: average number of products within a basket
```{r}

users <- orders %>%
  filter(eval_set == "prior") %>%
  group_by(user_id) %>%
  summarise(
    user_orders = max(order_number), #max order number
    user_period = sum(days_since_prior_order, na.rm = T), #days since first order
    user_mean_days_since_prior = mean(days_since_prior_order, na.rm = T) #average time between orders
  )

us <- orders_products %>%
  group_by(user_id) %>%
  summarise(
    user_total_products = n(),
    #user_reorder_ratio = sum(reordered == 1) / sum(order_number > 1),
    user_reorder_ratio = sum(reordered == 1) / n(),
    user_distinct_products = n_distinct(product_id)
  )

users <- users %>% inner_join(us)
users$user_average_basket <- users$user_total_products / users$user_orders

#identifier for users which are test and train users. 
us <- orders %>%
  filter(eval_set != "prior") %>%
  select(user_id, order_id, eval_set,
         time_since_last_order = days_since_prior_order)

users <- users %>% inner_join(us) #kritisch!
vh(users)

#order_id and eval_set are the identifier for train and test users last order. 
```


### Aisle and Department Features

* How often a specific department is bought by a user


# Create training dataset:

User & product specific features:
* up_orders: how often one specific user ordered a products
* up_first_order: when did this product occured at first
* up_last_order: when did the product occured at least
* up_average_cart_position: the average of when the product is added to the cart
* up_order_rate: % of how often a specific product was bought
* up_orders_since_last_order: difference between number of orders and the order when the product occured the last time
* up_order_rate_since_first_order: difference between number of orders and the order number when the product occured the first time. 

```{r}

data <- orders_products %>%
  group_by(user_id, product_id) %>% 
  summarise(
    up_orders = n(),
    up_first_order = min(order_number),
    up_last_order = max(order_number),
    up_average_cart_position = mean(add_to_cart_order))

head(data)

#join new features about user and products
data <- data %>% 
  inner_join(prd, by = "product_id") %>%
  inner_join(users, by = "user_id")

data$up_order_rate <- data$up_orders / data$user_orders
data$up_orders_since_last_order <- data$user_orders - data$up_last_order
data$up_order_rate_since_first_order <- 
  data$up_orders / (data$user_orders - data$up_first_order + 1)

#join column reordered (lable variable)
data <- data %>% 
  left_join(op_train %>% select(user_id, product_id, reordered), 
            by = c("user_id", "product_id"))

sum(is.na(data$reordered)) # not reordered
data %>% filter (reordered < 1) # contains just 1 for reordered

vh(data)

```

Here is the opportunity to scale some features:

```{r}
#data[,c(2:5,7,8,11)] <- scale(data[,c(2:5,7,8,11)])
#head(data)
```




Split data into train and test dataset:
```{r}
train <- as.data.frame(data[data$eval_set == "train",])

#added a split on the train to create a validationset
set.seed(123)
usersplit <- sample(unique(train$user_id),30000)
test <- train[train$user_id %in% usersplit,]
train <- train[!train$user_id %in% usersplit,]

#remove some columns
train$eval_set <- NULL
train$user_id <- NULL
#train$product_id <- NULL
train$order_id <- NULL
train$reordered[is.na(train$reordered)] <- 0 #reordered Nas == 0
#save(train, file = "train_xboost_vera_nuemro3.rda")

#test <- as.data.frame(data[data$eval_set == "test",]) #not used for validation
test$eval_set <- NULL
test$user_id <- NULL
test$reordered <- NULL
#save(test, file = "test_xboost_vera_numer03.rda")
```

# CV to find the best params:
can be fined in xgboost_vera_numero1_cv.R

# Fit model
```{r}
#new params:
#best_params <- read_csv("~/bigdataproject/03_XGBoost/2107xgboost_vera_best_parameters_FinalCV.csv")
#best_params

params <- list(
  "objective"           = "binary:logistic",
  "eval_metric"         = "logloss",
  "eta"                 = 0.1139337,
  "gamma"               = 0.1440471,
  "max_depth"           = 8,
  "min_child_weight"    = 17,
  "max_delta_step"      = 2, 
  "subsample"           = 0.718962,
  "colsample_bytree"    = 0.7965357,
  "alpha"               = 5.206276e-05,
  "lambda"              = 10
)

```

```{r}

subtrain <- train %>% sample_frac(0.5) #train just on a fraction due to computation probs
X <- xgb.DMatrix(as.matrix(subtrain %>% select(-reordered)), label = subtrain$reordered) #create xgb matrix
model <- xgboost(data = X, 
                 params = params, 
                 nrounds = 300,
                 nthread = 16
                 )

importance <- xgb.importance(colnames(X), model = model)
xgb.ggplot.importance(importance)


#ggsave(xgb.ggplot.importance(importance), file = "feature_importance.png")
```


# Predict on model: 
```{r}
Y <- xgb.DMatrix(as.matrix(test %>% select(-order_id)))
test$reordered <- predict(model, Y)

forcuts <- test %>% select(order_id,product_id,reordered)
save(forcuts,file="forcuts.rda")
#save(test, file= "2207predictions_xgboost_vera_numero3_newparams.rda")
#vh(test)
```

```{r}
test$reordered <- (test$reordered > 0.19) * 1
```

# Submission:
```{r}
submission <- test %>%
  filter(reordered == 1) %>% #select just products which are predicted as reordered
  group_by(order_id) %>%
  summarise(
    products = paste(product_id, collapse = " ")
  )

#some missing orders (75000 - 71044)
missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
  products = "None"
)

submission <- submission %>% bind_rows(missing) %>% arrange(order_id)
write.csv(submission, file = "xgboost_numero3_submission_2_new_params.csv", row.names = F)
```
