---
title: "F1 Testing"
author: "MP"
date: "1 7 2017"
output: html_document
---

In this code we will test different models for accuracy, using the F1 Score, which is defined in the chunk below. source: https://www.kaggle.com/rafaelmdourado/f1-score-with-r/notebook in the comments by grandmaster Konrad Banachewicz.

```{r F1 Score, message=FALSE, warning=FALSE}
require(stringr)
f1score <- function(list_a, list_b)
{
  list_a <- str_split(list_a, ' ')[[1]]
  list_b <- str_split(list_b, ' ')[[1]]
  x <- length(intersect(list_a, list_b))
  pr <- x/length(list_b)
  re <- x/length(list_a)
  f1 <- 0
  if (pr + re)
  {
    f1 <- 2 * pr * re /(pr  +re)
  }
  return(f1)
}
```


```{r Loading Data}
library(arulesViz)
library(dplyr)
library(markovchain)
load("/home/Vera_Weidmann/Supermarket/00_Data/boing.rda")
load("/home/Vera_Weidmann/Supermarket/00_Data/test_orders.rda")
```

here we compute the overall average F1 value, if we would predict the Basket t with Basket t-1

```{r global F1}
boing <- boing %>% group_by(user_id,order_number) %>% mutate(F1=f1score(vector1,vector2))
mean(boing$F1)
```
The mean F1 score is .17, which is below all kaggle competitors. lets check this again for only train users and there last basket versus the train-basket

```{r subset boing_train}
boing_train <- boing[!boing$user_id %in% test_orders$user_id,]
#save(boing_train, file="/home/Vera_Weidmann/Supermarket/00_Data/boing_train.rda")
```

```{r F1 train}
tmp <-boing_train %>% group_by(user_id) %>% filter(order_number==max(order_number))

tmp <- tmp %>% mutate(F1=f1score(vector1,vector2))

mean(tmp$F1)
```

Now we have a baseline, which is around 0.17, all our models and ideas which perform better than 0.17 in f1scores are cosidered valueable :)

```{r longdatasubset}
row=1
x <-cbind(rep(boing_train$vector1[[row]], each=length(boing_train$vector2[[row]])),rep(boing_train$vector2[[row]],length(boing_train$vector1[[row]])), rep(boing_train$user_id[row],length.out=length(boing_train$vector2[[row]])*length(boing_train$vector1[[row]])))

for(row in 2:114){
  y <-cbind(rep(boing_train$vector1[[row]], each=length(boing_train$vector2[[row]])),rep(boing_train$vector2[[row]],length(boing_train$vector1[[row]])), rep(boing_train$user_id[row],length.out=length(boing_train$vector2[[row]])*length(boing_train$vector1[[row]])))
  x <- rbind(x,y)
}
x <- as.data.frame(x)
```

```{r apriori}
baskets <- boing_train %>% group_by(user_id) %>% filter(order_number==max(order_number)) %>% ungroup() %>% select(vector2)
baskets <- c(boing_train$vector1, baskets$vector2)

allitems <- as.data.frame(as.character(unique(unlist(baskets)))) #all unique items
colnames(allitems) <- "items" 

baskets <- as(baskets, "transactions")

rules <- apriori(baskets,parameter=list(supp=0.00001, conf=0.5,target="rules",maxtime=30)) #create apriori rules, confidence 50% as threshold
```

replace x with longdata_train once it can be loaded

```{r MC-prep}
load("/home/Vera_Weidmann/Supermarket/00_Data/df_train_markov.rda")
actuals <- df_train_markov %>% 
  group_by(user_id) %>%
  summarise(basket=paste(product_id, collapse=" ")) %>%
  mutate(vector1= basket %>%
  strsplit(split = " ")) #get train in boing format

final_orders <- boing_train %>% group_by(user_id) %>% filter(order_number==max(order_number)) %>% select(user_id,basket, vector1) #select latest prior basket
```

```{r usermeans}
#whats the mean n of user?
#user_means <- boing_train %>% group_by(user_id,order_number) %>% summarise(n=length(vector1)) %>% summarise(m=mean(n)) %>% round(0)
```

```{r}
old <- array()
new <- array()
apr <- array()
tot <- array()
i=1
for(user in unique(x$V3)){
TransMC <- as.data.frame(markovchainFit(x %>% filter(V3==user) %>%select(V1,V2))$estimate@transitionMatrix) #mini markovchain
last <- final_orders[final_orders$user_id==user,3]
actual <- actuals[actuals$user_id==user,3] # train

old[i] <- f1score(last,actual)

last <- unlist(last) # last prior
preds <- TransMC[last,] %>% colMeans() %>% sort(decreasing=TRUE) #predictions on mini mc

preds.names <- list(c(names(preds[preds>0]))) # cut preds
new[i] <- f1score(preds.names,actual)
preds.names <- names(preds[preds>0]) # cut preds

#need workaround in case rulesubset is not existing
rules.sub <- subset(rules, subset = !lhs %in% as.character(allitems[!allitems$items %in% preds.names,1]))

rhs.desc <- as(inspect(rules.sub), "data.frame")[,-2] %>% arrange(-confidence) %>% mutate(rhs=gsub("[{}]","",rhs))
#better algorithm from here
rhs <- unique(rhs.desc$rhs)
apr[i] <- f1score(list(rhs),actual)
newbasket <- unique(c(rhs,preds.names))

tot[i] <- f1score(list(newbasket),actual)
i =i+1
}

mean(old);mean(new);mean(apr);mean(tot)
```

