---
title: "Supermarket Basket Analysis"
author: "Vera"
date: "6 6 2017"
output: html_document
---

This script is for the markov chain analysis

Notes for later:
* combine sales from same day, same user
* predict how many products will be bought by using predictors day, hour and new feature: how many products had been in the last purchase

Libraries:
```{r, message=FALSE}
library(dplyr)
library(reshape2)
```

Load Data:
-> For Changing or creating new rda files, pls, use the following script: "Script Loading Data.rmd"
```{r}
load("/home/Vera_Weidmann/Supermarket/00_Data/df_markov.rda") #gemerged
load("/home/Vera_Weidmann/Supermarket/00_Data/df_vali_markov.rda") #train
load("/home/Vera_Weidmann/Supermarket/00_Data/df_train_markov.rda") #prior

df_train_markov <- df_markov %>%
  arrange(user_id, order_number)
df_vali_markov <- df_markov %>%
  arrange(user_id, order_number)
df_markov <- df_markov %>%
  arrange(user_id, order_number)
```


#Non-personalized Markov chain

One-next-step-thingy
```{r}
#subset for user1:
#df_user1 <- df_markov %>%
 # filter (user_id == 1)

#save(df_user1, file = "/home/Vera_Weidmann/Supermarket/00_Data/df_user1.rda")
```


creation of transition matrix
```{r}
#set the data to name x
data <- df_train_markov
df <- as.data.frame(matrix(ncol=2))

for (i in unique(data$user_id)){
  x <- data %>% filter(user_id==i)
  for (j in unique(x$order_number)) {
    if (j ==max(x$order_number)) {break}
    
    vector1 <- x$product_id[x$order_number==j]
    vector2 <- x$product_id[x$order_number==j+1]
    
    tmp <- cbind(rep(vector1,each=length(vector2)), rep(vector2,length(vector1)))
    df <- rbind(df,tmp)
    print(paste(i,j,sep=" "))
  }
}


df <- df[-1,] %>% group_by(V1,V2) %>% summarise(n=n()) %>% mutate(p=n/max(n)) %>% dcast(V1 ~ V2, value.var="p")
df[is.na(df)] = 0
```

mean basketsize per customer, replace with model later

```{r}
#whats the mean n of user?
user_means <- data %>% group_by(user_id,order_number) %>% summarise(n=n()) %>% summarise(m=mean(n)) %>% round(0)
```

```{r}
#initializing global variables
results <- as.data.frame(matrix(ncol=2))
r <- 1
qual <- as.vector(0)


y <- df_vali_markov



for (i in unique(data$user_id)){
  trainvector <- data$product_id[(x$order_number==max(x$order_number) & x$user_id==i)] #last basket for user i in data
  
  predictions <- df[df$V1 %in% trainvector,-1] %>% colMeans() %>% sort(decreasing=TRUE)

  preds <- names(predictions[1:user_means$m[user_means$user_id==i]]) #cutting predictions at the estimated length
  
  controlvector <- y$product_id[y$user_id==i] #validation set

  qual[r] <- mean(names(predictions[1:user_means$m[user_means$user_id==i]]) %in% controlvector) * 
          mean(!controlvector %in% names(predictions[1:user_means$m[user_means$user_id==i]]))
  
  
  results[r,1] <- i
  results[r,2] <- paste(preds,collapse=" ")

  r=r+1
}
hist(qual)
```
