---
title: "Supermarket Basket Analysis"
author: "Vera"
date: "6 6 2017"
output: html_document
---

This script is for the markov chain analysis

Notes for later:
* combine sales from same day, same user
* predict how many products will be bought by using predictors day, hour and new feature: how many products had been in the last purchase

Libraries:
```{r, message=FALSE}
library(dplyr)
library(reshape2)
```

Load Data:
-> For Changing or creating new rda files, pls, use the following script: "Script Loading Data.rmd"
```{r}
load("/home/Vera_Weidmann/Supermarket/00_Data/df_prior_markov.rda")
df_prior_markov <- df_prior_markov %>%
  arrange(user_id, order_number)
```

```{r}
#testing data for just 20 people
#df_prior_markov500 <- df_prior_markov %>%
#  filter (user_id %in% 1:500)
```

Dplyr only Matrix creation
```{r}

data <- df_prior_markov #train and test people


df <- as.data.frame(matrix(ncol=2))


boing <- data %>% 
  group_by(user_id,order_number) %>%
  summarise(basket=paste(product_id, collapse=" ")) %>%
  group_by(user_id) %>%
  mutate(vector1= basket %>%
           strsplit(split = " "), vector2=lead(basket) %>% strsplit(split = " ")) %>%
  ungroup() %>% 
  filter(!is.na(vector2)) 

save(boing, file = "/home/Vera_Weidmann/Supermarket/00_Data/boing.rda")

# for (i in 1:nrow(boing)){
#   df <- rbind(df,
#     cbind(rep(boing$vector1[[i]], each=length(boing$vector2[[i]])), rep(boing$vector2[[i]],length(boing$vector1[[i]]))))
# }
# 
# ptm <- proc.time()
# df3 <- df[-1,] %>% group_by(V1,V2) %>% summarise(n=n()) %>% mutate(p=n/max(n)) %>% dcast(V1 ~ V2, value.var="p")
# df3[is.na(df3)] = 0

```


#Non-personalized Markov chain

One-next-step-thingy
```{r}
#subset for user 1-5:
#df_user5_prior <- df_prior_markov_usertest %>%
#  filter (user_id %in% 1:20) %>%
#  arrange (user_id, order_number)

#df_user5_train <- df_train_markov_us %>%
#  filter (user_id %in% 1:5) %>%
 # arrange (user_id, order_number)

#save(df_user1, file = "/home/Vera_Weidmann/Supermarket/00_Data/df_user1.rda")
```


creation of transition matrix
```{r}
#set the data to name x
data <- df_prior_markov #train and test people
#data = df_user5_prior
df <- as.data.frame(matrix(ncol=2))

for (i in unique(data$user_id)){
  x <- data %>% filter(user_id==i)
  for (j in unique(x$order_number)) {
    if (j ==max(x$order_number)) {break}
    
    vector1 <- x$product_id[x$order_number==j]
    vector2 <- x$product_id[x$order_number==j+1]
    
    tmp <- cbind(rep(vector1,each=length(vector2)), rep(vector2,length(vector1)))
    df <- rbind(df,tmp)
    #print(paste(i,j,sep=" "))
  }
}

df <- df[-1,] %>% group_by(V1,V2) %>% summarise(n=n()) %>% mutate(p=n/max(n)) %>% dcast(V1 ~ V2, value.var="p")
df[is.na(df)] = 0
```

mean basketsize per customer, replace with model later

```{r}
#whats the mean n of user?
user_means <- data %>% group_by(user_id,order_number) %>% summarise(n=n()) %>% summarise(m=mean(n)) %>% round(0)
```

```{r}
#initializing global variables
results <- as.data.frame(matrix(ncol=2))
r <- 1
qual <- as.vector(0)

#y <- df_vali_markov
#y <- df_user5_train


for (i in unique(data$user_id)){
  x <- data %>% filter(user_id==i)
  trainvector <- x$product_id[(x$order_number==max(x$order_number))] #last basket for user i in data
  predictions <- df[df$V1 %in% trainvector,-1] %>% colMeans() %>% sort(decreasing=TRUE)

  preds <- names(predictions[1:user_means$m[user_means$user_id==i]]) #cutting predictions at the estimated length
  
  #controlvector <- y$product_id[y$user_id==i] #validation set

  #qual[r] <- mean(names(predictions[1:user_means$m[user_means$user_id==i]]) %in% controlvector) * 
 #         mean(!controlvector %in% names(predictions[1:user_means$m[user_means$user_id==i]]))
  
  
  results[r,1] <- i
  results[r,2] <- paste(preds,collapse=" ")

  r=r+1
}
#hist(qual)
```
