---
title: "testing cutoffs"
author: "MP"
date: "18 7 2017"
output: html_document
---
```{r}
load("/home/Max_Philipp/bigdata/cutofflookup.rda")
load("/home/Vera_Weidmann/bigdataproject/00_vera_dont_push/save_predictions_xgboost_vera_numero2_for_max.rda")
lookup <- lookup %>% group_by(user_id) %>% mutate(cutoff=max(round((user_reorder_ratio+(1-user_reorder_ratio)/2)*user_average_basket),1))
library(dplyr)
```

```{r}
hist(test$reordered)

```
now i will test the leaderboard score on multiple cutoff ideas to see if there is one particular which outperforms all the other.

first idea: inverse reorder ratio. If an user has a reorderratio of 80%, he only needs 1-0.8=20% security to buy again. - 0.2677424 LBS

```{r}
test$reordered <- (test$reordered > 0.18) * 1

sub.pred <- test %>% select(order_id,product_id, reordered) %>% left_join(lookup, by="order_id") %>% mutate(invratio=1-user_reorder_ratio) #valid for all methods

inv.pred <- sub.pred %>% group_by(order_id) %>% filter(reordered>=invratio) %>% summarise(products = paste(product_id, collapse = " "))
missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% inv.pred$order_id]),
  products = "None"
)
submission <- inv.pred %>% bind_rows(missing) %>% arrange(order_id)
write.csv(submission, file = "sub_inverseratio.csv", row.names = F)

inv.pred <- sub.pred %>% group_by(order_id) %>% mutate(reordered=ifelse(reordered>=invratio,1,0))

```

second idea: cutoff on average basket size per user -0.3749070 LSB
```{r}

submission <- sub.pred %>%
  arrange(user_id,-reordered)%>% 
  group_by(user_id) %>% 
  mutate(index=row_number()) %>% 
  filter(index<=user_average_basket) %>% 
  ungroup() %>% 
  group_by(order_id) %>% 
  summarise(products = paste(product_id, collapse = " "))
write.csv(submission, file = "sub_mean_cutoff.csv", row.names = F)

mean.pred <- sub.pred %>%
  arrange(user_id,-reordered)%>% 
  group_by(user_id) %>% 
  mutate(index=row_number(),reordered=ifelse(index<=user_average_basket,1,0))

```

idea three: using a modified cutoff, representing that a mean basket of 10 and a reorder ratio of .6 only explains 6/10 of the basket, leaving 4/10 unexplained. - 0.3733417 LSB
```{r}
submission <- sub.pred %>%
  arrange(user_id,-reordered)%>% 
  group_by(user_id) %>% 
  mutate(index=row_number()) %>% 
  filter(index<=cutoff) %>% 
  ungroup() %>% 
  group_by(order_id) %>% 
  summarise(products = paste(product_id, collapse = " "))
write.csv(submission, file = "sub_enhanced_mean_cutoff.csv", row.names = F)

enh.mean.pred <- sub.pred %>%
  arrange(user_id,-reordered)%>% 
  group_by(user_id) %>% 
  mutate(index=row_number(),reordered=ifelse(index<=cutoff,1,0))
```

idea four: using a global cutoff. testing on usefullness with cutoff 0.18 - 0.3799640 LBS
```{r}
global.pred <- sub.pred %>% mutate(reordered=reordered > 0.18 * 1)
submission <- global.pred %>%
  filter(reordered == 1) %>%
  group_by(order_id) %>%
  summarise(
    products = paste(product_id, collapse = " ")
  )

missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
  products = "None"
)
submission <- submission %>% bind_rows(missing) %>% arrange(order_id)
write.csv(submission, file = "sub_global_18.csv", row.names = F)
```

0.3813866
```{r}
global.pred <- sub.pred %>% mutate(reordered=reordered > 0.21 * 1)
submission <- global.pred %>%
  filter(reordered == 1) %>%
  group_by(order_id) %>%
  summarise(
    products = paste(product_id, collapse = " ")
  )

missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
  products = "None"
)
submission <- submission %>% bind_rows(missing) %>% arrange(order_id)
write.csv(submission, file = "sub_global_21.csv", row.names = F)
```

idea five: majority vote among all 4 methods
```{r}
all.pred <- mean.pred[,c(1:3)] %>%
  left_join(enh.mean.pred[,c(1:3)], 
            by=c("order_id","product_id")) %>%
  left_join(inv.pred[,c(1:3)], 
            by=c("order_id","product_id")) %>%
  left_join(global.pred[,c(1:3)], 
            by=c("order_id","product_id")) %>%
  rename(meanbasket=reordered.x,
         enhanced=reordered.y,
         inverted=reordered.x.x,
         global_21=reordered.y.y) %>% 
  mutate(global_21=as.numeric(global_21))

save(all.pred, file="concat_preds_model2.rda")
load("concat_preds_model2.rda")
```

